---
title: "sf"
author: "Joseph Adams"
date: "14 April 2019"
output: html_document
---

These examples follow this `sf` [vignette](https://r-spatial.github.io/sf/articles/sf1.html)

```{r message = FALSE}
library(ggplot2)
library(sf)
library(dplyr)
```

`sf` comes with some built in datasets:

```{r}
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
```

`nc` is an object of class `sf` *and* `data.frame`. So it is a dataframe but also extends the dataframe class. The extra bit you get is the geom column of class `sfc`.

```{r}
class(nc)
```

```{r}
attr(nc, "sf_column")
```

```{r}
print(nc[9:15], n = 3)
```

Retrieve the geometry list-column with `st_geometry`

```{r}
(nc_geom <- st_geometry(nc))
```

Here's the first complete geometry in the geometry list-column

```{r}
nc_geom[[1]]
```

```{r}
plot(nc[1], reset = FALSE)
plot(nc[1,1], col = 'white', add = TRUE)
```

Only plot the polygons with multiple rings.
```{r}
par(mar = c(0,0,1,0))
(w <- which(sapply(nc_geom, length) > 1))
## [1]  4 56 57 87 91 95
plot(nc[w,1], col = 2:7)
```

The geometry column has it's own special class

```{r}
class(nc_geom)
```

A tweaked version of the ggplot2 example in this `sf` vignette on [plotting simple features](https://r-spatial.github.io/sf/articles/sf5.html#ggplot2)

```{r}
# Download shapefile from:
# https://geoportal.statistics.gov.uk/datasets/local-authority-districts-december-2018-boundaries-uk-bgc

la <- st_read("sf_files/Local_Authority_Districts_December_2018_Boundaries_UK_BGC.shp",
              quiet = TRUE)

mw <- read.csv("https://statistics.gov.scot/downloads/cube-table?uri=http%3A%2F%2Fstatistics.gov.scot%2Fdata%2Fmunicipal-waste") %>%
  filter(Indicator..municipal.waste. == 'Waste Recycled') %>%
  filter(DateCode == '2011/2012') %>%
  filter(FeatureCode != 'S92000003')

la_mw <- merge(la, mw, by.x = 'lad18cd', by.y = 'FeatureCode', all.x = FALSE)
```

```{r}
ggplot() +
  geom_sf(data = la_mw["Value"], mapping = aes(fill = Value), colour = "white") +
  theme(
    # Declutter
    panel.background = ggplot2::element_blank(),
    legend.position = "none",
    axis.title = ggplot2::element_blank(),
    axis.ticks = element_blank(),
    
    # Text
    title = element_text(size = 13),
    axis.text = ggplot2::element_blank()
  )
```